# Raster-vector interactions

**Learning objectives:**

- Raster cropping/masking

- Extracting raster value with vector data

- Raster <-> vector conversion i

```{r}
library(sf)
library(terra)
library(dplyr)
```


## Raster cropping/masking

One (very) common problem:  the extent of input raster is larger than the area of interest. 

Solution: Cropping and masking!

```{r}
# a raster:  SpatRaster
srtm = terra::rast(system.file("raster/srtm.tif", package = "spDataLarge"))
zion = sf::read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
# same CRS for both of this object
zion = st_transform(zion, crs(srtm))
# always do a quick viz
plot(srtm, col = grDevices::terrain.colors(50))
plot(zion$geom, add = TRUE)
```

## Raster cropping/masking: Crop

We are limiting the extend: 

```{r}
srtm_cropped = terra::crop(srtm, zion)
plot(srtm_cropped, col = grDevices::terrain.colors(50))
plot(zion$geom, add = TRUE)
```

## Raster cropping/masking: mask

Value outside of our area of interest: `NA` (could be change with `updatevalue`).

```{r}
srtm_masked = terra::mask(srtm, zion)
plot(srtm_masked, col = grDevices::terrain.colors(50))
plot(zion$geom, add = TRUE)
# srtm_inv_masked = mask(srtm, zion, inverse = TRUE)
```

## Raster extraction

We want to extract values from "target" based on a selector (lot of time a vector). The result will depend on the type of object (point, lines, polygons) we are using to select.

### Extracting value on points

```{r}
data("zion_points", package = "spDataLarge")
elevation = terra::extract(srtm, zion_points)
zion_points_elev = cbind(zion_points, elevation)

plot(srtm_masked, col = grDevices::terrain.colors(50))
plot(zion_points, add = TRUE, pch = 16)
```

### Extracting value on lines

First we create a line:

```{r}
zion_transect = cbind(c(-113.2, -112.9), c(37.45, 37.2)) |>
  sf::st_linestring() |> 
  sf::st_sfc(crs = crs(srtm)) |>
  sf::st_sf(geometry = _ ) # <- fancy new place holder

plot(srtm_masked, col = grDevices::terrain.colors(50))
plot(zion_transect, add = TRUE, lwd = 3)
```

The we are using points on this line: 

```{r}
zion_transect$id = 1:nrow(zion_transect) # just on transect
zion_transect = st_segmentize(zion_transect, dfMaxLength = 250) # I am curious why I get 257 pts
# we are casting as POINT
zion_transect = st_cast(zion_transect, "POINT")
```

 I am not sure we need the `group_by()` ?

```{r}
zion_transect = zion_transect |> 
  #group_by(id) |> 
  mutate(dist = st_distance(geometry)[, 1]) 
```

```{r}
zion_elev = terra::extract(srtm, zion_transect)
# just cbind 
zion_transect = cbind(zion_transect, zion_elev)
# quick plot
plot(zion_transect$dist, zion_transect$srtm, type = "l")
```


### Extracting value on polygons

```{r}
zion_srtm_values = terra::extract(x = srtm, y = zion)
dim(zion_srtm_values)
```

### With quantitative data:

```{r}
group_by(zion_srtm_values, ID) |> 
  summarize(across(srtm, list(min = min, mean = mean, max = max)))
```

### With qualitative data:

```{r}
nlcd = rast(system.file("raster/nlcd.tif", package = "spDataLarge"))
zion2 = st_transform(zion, st_crs(nlcd))
zion_nlcd = terra::extract(nlcd, zion2)
zion_nlcd |> 
  group_by(ID, levels) |>
  count()
```

**Important!** If you are using a lot of polygons you should use [exactextract](https://github.com/isciences/exactextractr) 

## Rasterization

Vector ---> Raster : `terra::rasterize()`

We are using a data set from `spData`:

```{r}
# the easy part
cycle_hire_osm = spData::cycle_hire_osm
cycle_hire_osm_projected = st_transform(cycle_hire_osm, "EPSG:27700")
# The new part: building a raster template
raster_template = terra::rast(ext(cycle_hire_osm_projected), resolution = 1000,
                       crs = st_crs(cycle_hire_osm_projected)$wkt)
```

- First Presence/absence:

```{r}
ch_raster1 = rasterize(cycle_hire_osm_projected, raster_template)
plot(ch_raster1, col = "yellow")
```

- Second: count with `length` :

```{r}
ch_raster2 = rasterize(cycle_hire_osm_projected, raster_template, 
                       fun = "length")
plot(ch_raster2)
```


- Third: sum of a variable

```{r}
ch_raster3 = terra::rasterize(cycle_hire_osm_projected, raster_template, 
                       field = "capacity", fun = sum, na.rm = TRUE)
plot(ch_raster3)
```

`rasterize()` has an argument `touches` that is useful when dealing with lines or polygon border.


## Spatial vectorization

raster ---> vector: `as.point s()`, `as.polygon()`

```{r}
elev = rast(system.file("raster/elev.tif", package = "spData"))
elev_point = terra::as.points(elev) |> 
  st_as_sf()
plot(elev_point["elev"])
```

```{r}
dem = rast(system.file("raster/dem.tif", package = "spDataLarge"))
cl = terra::as.contour(dem) |> 
  st_as_sf()
plot(dem, axes = FALSE)
plot(cl, add = TRUE)
```

```{r}
grain = rast(system.file("raster/grain.tif", package = "spData"))
grain_poly = terra::as.polygons(grain) |> 
  st_as_sf()
plot(grain_poly)
```

```{r}
grain_poly = terra::as.polygons(grain, dissolve = FALSE) |> 
  st_as_sf()
plot(grain_poly)
```


## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/iUUxI4B_zKE")`

<details>
<summary> Meeting chat log </summary>

```
00:26:13	Derek Sollberger:	srtm == "shuttle radar topography mission"?  Is that right?
00:26:26	Tony Vota:	I think that is right
00:26:32	Tony Vota:	he Shuttle Radar Topography Mission (SRTM) uses inSAR which measures Earth's elevation with two antennas. Using two radar antennas and a single pass, it generated a digital elevation model using interferometric synthetic aperture radar (InSAR).
01:01:30	Derek Sollberger:	The presentation of chapter 7 perhaps should be handled by someone who works with multiple projection systems regularly
01:03:20	Tony Vota:	Yes, thank you.
01:03:37	Tony Vota:	Liked how you show each step of the piped transactions
```
</details>
