# Reprojecting geographic data

Build on previous chapters: *geographic*/*projected* CRS 

**Learning objectives:**

- Interact with CRS (`sf::st_crs`, `terra::crs`)
- What kind of trouble we can run into misusing CRS
- When/how to transform  (`sf::st_transform`, `terra::project`) 
- Which one to use   

```{r}
library(sf)
library(terra)
library(dplyr)
library(spData)
library(spDataLarge)
```


## Coordinate Reference System (CRS)

Upstream: [PROJ](https://proj.org/development/quickstart.html)

Sometimes CRS can be loosely defined: 

- "lat/long"   

- Outdated *proj4string* (just *projstring*):  `+proj=longlat +ellps=WGS84 +datum=WGS84 +no_def`  
- "AUTHORITY:CODE": `EPSG:4326`  

They are the same but only the last one is rigorous and less ambiguous. 

We can inspect them that way: 

```{r}
sf::st_crs("EPSG:4326")
# investigated the R object
```

-> https://epsg.io/

We are currently using WKT2 (but I see regularly WKT1 and proj4string). 

The last line is also called **SRID**: Spatial Reference IDentifier. 

## Querying and Setting CRS

### For vector:

- `sf::st_crs` : get the CRS

- `sf::st_crs("ESPG::4326")$input` : how the user specified the CRS

- `sf::st_crs("ESPG:4326")$wkt` : WKT, use `cat()`

```{r}
sf::st_set_crs # is a convenient function
```

### Raster:

```{r, eval=FALSE}
terra::crs(my_rast) = "EPSG:26912"
```


**Attention!**: Setting does not alter coordinates values or geometries

You can also directly specify CRS with `crs` argument in `sf::st_crs()`/ `sf::st_as_asf()`  and `terra::rast()`

## Geometry operations on projected/unprojected data

sf use GEOS for projected and s2 for geographic CRS. 

```{r}
london = data.frame(lon = -0.1, lat = 51.5) |> 
  st_as_sf(coords = c("lon", "lat"))
st_is_longlat(london)
london_geo = st_set_crs(london, "EPSG:4326")
st_is_longlat(london_geo)
```

```{r}
london_buff_no_crs = st_buffer(london, dist = 1)   # incorrect: no CRS
london_buff_s2 = st_buffer(london_geo, dist = 100000) # silent use of s2
london_buff_s2_100_cells = st_buffer(london_geo, dist = 100000, max_cells = 10000) # play with max_cells 
# sf::sf_use_s2(FALSE)
london_proj = data.frame(x = 530000, y = 180000) |> 
  st_as_sf(coords = c("x", "y"), crs = "EPSG:27700")
london_buff_projected = st_buffer(london_proj, 100000)
```


In geographic CRS:

```{r}
UK <- world[world$name_long == "United Kingdom",]
plot(UK$geom)
plot(london_buff_s2_100_cells, col = "red", add = TRUE)
plot(london_buff_no_crs$geometry, add = TRUE, lwd = 3)
```

In projected CRS:

```{r}
plot(st_transform(UK$geom, "EPSG:27700"))
plot(london_buff_projected$geometry, add = TRUE, lwd = 3, col = "blue")
```


## When to project?

Hard to provide a generic answers but:

- if you want to use leaflet: "EPSG:4326"  

- When you want to compare two objects (or) they need to be in the same CRS

## Which CRS to use?

Same: no good answers! 

-> Use the right one for every task

If it is a geographic: the default is "EPSG:4326"

### Projected:

- Use the one that the public mapping agency use (In France "EPSG:2154")

- UTM: for relatively small area (less than 6 degrees in longitude), you will have a distorted area farther you go from the center of UTM.

- You can use https://crs-explorer.proj.org/


You can pick: 

- Conformal: preserving angles -> same shapes 

- Equal area: keep area the same: good for density

- equidistant: preserving distance

- a mix of the above 

The goal will be to minimize distortion. 

Authors recommendations: 

- Lambert Azimuthal Equal Area ([LAEA](https://en.wikipedia.org/wiki/Lambert_azimuthal_equal-area_projection)): same size but distorted shape

- Azimuthal equidistant ([AEQD]https://en.wikipedia.org/wiki/Azimuthal_equidistant_projection)): accurate line distance between a point and the center point of the local projection

- Lambert Conformal Conic ([LCC](https://en.wikipedia.org/wiki/Lambert_conformal_conic_projection)): good if you are away of equator

- Stereographic ([STERE](https://en.wikipedia.org/wiki/Stereographic_map_projection)

## Reprojecting vector geometries:

Transforming coordinate points (vertices) and usually setting CRS

```{r}
london2 = st_transform(london_geo, "EPSG:27700")
```


```{r}
crs_lnd = st_crs(london_geo) # see with london_proj
class(crs_lnd) # we already see that
names(crs_lnd) # same
crs_lnd$Name
crs_lnd$proj4string
crs_lnd$epsg
```

## Reprojecting raster:

- create a new raster object (with different rows/cols)

- resampling values of every cell (using different resamploing methods)  

Terra's function `project()` allow you to do that.

As usual: nearest neighbour method  are good for categorical data. You have more option for quantitative data.

```{r}
cat_raster = rast(system.file("raster/nlcd.tif", package = "spDataLarge"))
crs(cat_raster) |> cat()
```

```{r}
cat_raster_wgs84 = project(cat_raster, "EPSG:4326", method = "near")
cat_raster_wgs84
cat_raster
```

## My tl:dr

- Check good references before doing anything funny

- Use recommended CRS by mapping institutions  

- on small area it does not mapper too much


## Resources:

- http://www.geo.hunter.cuny.edu/~jochen/gtech201/lectures/lec6concepts/map%20coordinate%20systems/how%20to%20choose%20a%20projection.htm#

- https://ihatecoordinatesystems.com/

- https://mgimond.github.io/tissot/

## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/RuvLbpyrDmU")`

<details>
<summary> Meeting chat log </summary>

```
00:17:17	Derek Sollberger:	https://epsg.io/4326
00:28:10	Derek Sollberger:	Is S2 a type of object-oriented programming?
00:36:35	Derek Sollberger:	https://crs-explorer.proj.org/
00:36:54	Jim Gruman:	Reacted to "https://crs-explorer..." with 👍🏼
00:44:36	Derek Sollberger:	I have seen "Mollweide" in a few places for displaying the entire Earth.  What is the motivation for that projection?
```
</details>
