# Geomarketing

People are fundamental to location analysis, in particular where they are likely to spend their time and other resources.

Humans tend to congregate in certain places, creating economic niches (and high land prices).

The main task of location analysis is to find out, based on available data, where such ‘optimal locations’ are for specific services.

Typical research questions include:

- Where do target groups live and which areas do they frequent?
- Where are competing stores or services located?
- How many people can easily reach specific stores?
- Do existing services over- or under-utilize the market potential?
- What is the market share of a company in a specific area?

```{r}
#| label: 14 setup

library(sf)
library(dplyr)
library(ggplot2)
library(purrr)
library(terra)
library(osmdata)
library(spDataLarge)
library(tmap)
library(mapview)

tmap_mode("plot")
```

## Case Study: Bike Shops in Germany

We are starting a new chain, near as many potential customers as possible.

Our target audience must include males age 20-40.

Where should the stores be placed?

## Tidy input data

The German government provides gridded census data at either 1 km or 100 m resolution. The following code chunk downloads, unzips and reads in the 1 km data.

```{r}
#| label: 14 input data

download.file("https://tinyurl.com/ybtpkwxz", 
              destfile = "census.zip", mode = "wb")

unzip("census.zip") # unzip the files

input_tidy <-
  readr::read_csv2(list.files(pattern = "Gitter.csv")) |>
  select(
    x = x_mp_1km,
    y = y_mp_1km,
    pop = Einwohner,
    women = Frauen_A,
    mean_age = Alter_D,
    hh_size = HHGroesse_D
  ) |>
  dplyr::mutate(dplyr::across(
                  .cols = c(pop, women, mean_age, hh_size),
                  .fns =  ~ ifelse(.x %in% c(-1,-9), NA, .x)
                ))

summary(input_tidy)
```

## Create census rasters

The `x` and `y` from the input file correspond to a 1000 x 1000 square meter unit grid.

```{r}
#| label: 14 create rasters

(input_ras <- terra::rast(input_tidy, type = "xyz", crs = "EPSG:3035"))

ger <- geodata::gadm(country = "DEU", level = 0, path = tempdir())

ger <- st_as_sf(terra::project(ger, crs(input_ras)))

tm_shape(input_ras) +
  tm_raster(style = "cat", palette = "GnBu", title = "Class") +
  tm_facets(nrow = 1) +
  tm_shape(ger) +
  tm_borders() +
  tm_layout(panel.labels = c("population", "women", "mean age", "household size"),
            legend.outside.size = 0.08)
```

The next stage is to reclassify the values of the rasters stored in input_ras in accordance with the survey mentioned above, using the terra function `classify()`

In the case of the population data, we convert the classes into a numeric data type using class means. 

Raster cells are assumed to have a population of 127 if they have a value of 1 (the middle of the categorical bin)

The remaining variables are re-classified as weights corresponding with weights used in the survey. 

Class 1 in the variable women represents areas in which 0 to 40% of the population is female; these are reclassified with a comparatively high weight of 3 because the target demographic is predominantly male. Similarly, the classes containing the youngest people and highest proportion of single households are reclassified to have high weights.

```{r}
#| label: 14 weightings

rcl_pop <- matrix(c(1, 1, 127, 2, 2, 375, 3, 3, 1250, 
                   4, 4, 3000, 5, 5, 6000, 6, 6, 8000), 
                 ncol = 3, byrow = TRUE)

rcl_women <- matrix(c(1, 1, 3, 2, 2, 2, 3, 3, 1, 4, 5, 0), 
                   ncol = 3, byrow = TRUE)

rcl_age <- matrix(c(1, 1, 3, 2, 2, 0, 3, 5, 0),
                 ncol = 3, byrow = TRUE)

rcl_hh <- rcl_women

rcl <- list(rcl_pop, rcl_women, rcl_age, rcl_hh)

reclass <- input_ras

for (i in seq_len(terra::nlyr(reclass))) {
  reclass[[i]] = terra::classify(x = reclass[[i]], rcl = rcl[[i]], right = NA)
}

names(reclass) <- names(input_ras)

reclass
```

## Define metropolitan areas

We define metropolitan areas here as pixels of 20 km2 inhabited by more than 500,000 people


```{r}
#| label: 14 Urban

pop_agg <- terra::aggregate(reclass$pop, fact = 20, fun = sum, na.rm = TRUE)

pop_agg <- pop_agg[pop_agg > 500000, drop = FALSE] 

summary(pop_agg)

metros <- pop_agg |> 
  terra::patches(directions = 8) |>
  terra::as.polygons() |>
  sf::st_as_sf()

metros$names = c("Hamburg", "Berlin", "Düsseldorf", "Leipzig",
                 "Frankfurt am Main", "Nürnberg", "Stuttgart", "München")

metros_points <- st_centroid(metros)

tm_shape(pop_agg/1000) +
  tm_raster(palette = "GnBu",
            title = "Number of people\n(in 1,000)") +
  tm_shape(ger) +
  tm_borders() +
  tm_shape(metros) +
  tm_borders(col = "gold", lwd = 2) +
  tm_shape(metros_points) +
  tm_text(text = "names", ymod = 0.6, shadow = TRUE, size = 0.75,
          fontface = "italic") +
  tm_layout(legend.outside = TRUE, legend.outside.size = 0.3)
```

We could also reverse geocode for the city names

```{r}
#| label: 14 reverse geocode

metro_names <- sf::st_centroid(metros, of_largest_polygon = TRUE) |>
  tmaptools::rev_geocode_OSM(as.data.frame = TRUE) |>
  select(city, town, state)

metro_names <- dplyr::mutate(metro_names, city = ifelse(is.na(city), town, city))

metro_names
```


## Points of interest

poi (Points of Interest)

The `osmdata` package provides easy-to-use access to Open Street Map data

Caution:  even filtering for metro areas in Germany, this query operates on 2Gb of data.

```{r eval=FALSE}
#| label: 14 osmdata import

shops <- purrr::map(metro_names, function(x) {
  message("Downloading shops of: ", x, "\n")

  Sys.sleep(sample(seq(5, 10, 0.1), 1))
  query = osmdata::opq(x) |>
    osmdata::add_osm_feature(key = "shop")
  points = osmdata::osmdata_sf(query)

  iter = 2
  while (nrow(points$osm_points) == 0 && iter > 0) {
    points = osmdata_sf(query)
    iter = iter - 1
  }

  points$osm_points
})

# checking if we have downloaded shops for each metropolitan area
ind = purrr::map_dbl(shops, nrow) == 0
if (any(ind)) {
  message("There are/is still (a) metropolitan area/s without any features:\n",
          paste(metro_names[ind], collapse = ", "), "\nPlease fix it!")
}

# select only specific columns
shops = purrr::map_dfr(shops, select, osm_id, shop)

```

We will just use the convenient sample dataset with features in each of the metro areas.

```{r}
#| label: 14 osmdata sample from spDataLarge

data("shops", package = "spDataLarge")

shops |> 
  ggplot() +
  geom_sf(alpha = 0.01, shape = 20) +
  theme_minimal()
```

This spatial point object must be converted into a raster 

```{r}
#| label: 14 convert shops to a raster

shops <- sf::st_transform(shops, st_crs(reclass))

poi <- terra::rasterize(x = shops, y = reclass, field = "osm_id", fun = "length")
```

As with the other raster layers (population, women, mean age, household size) the poi raster is reclassified into four classes.  Here again, some judgement is needed.

The authors choose the Fisher-Jenks natural breaks approach which minimizes within-class variance, the result of which provides an input for the reclassification matrix.

```{r}
#| label: 14 shops reclassification

int <- classInt::classIntervals(terra::values(poi), n = 4, style = "fisher")

int <- round(int$brks)

rcl_poi <- matrix(c(int[1], rep(int[-c(1, length(int))], each = 2),
                    int[length(int)] + 1), ncol = 2, byrow = TRUE)

rcl_poi <- cbind(rcl_poi, 0:3)  

poi <- terra::classify(poi, rcl = rcl_poi, right = NA) 

names(poi) = "poi"

poi
```

## Identifying suitable locations

Our last step: add `poi` to the reclass raster stack and remove the population layer from it. 

The reasoning: 
First of all, we have already delineated metropolitan areas, that is areas where the population density is above average compared to the rest of Germany. 

Second, though it is advantageous to have many potential customers within a specific catchment area, the sheer number alone might not actually represent the desired target group. For instance, residential tower blocks are areas with a high population density but not necessarily with a high purchasing power for expensive cycle components.

```{r}
#| label: 14 remove population raster and add poi 

reclass2 <- reclass[[names(reclass) != "pop"]] |>
  c(poi)

plot(reclass2)
```

the final step — calculating a final score by summing the scoring across all of the raster layers

```{r}
#| label: 14 sum of scores

(result <- sum(reclass2))

# have a look at suitable bike shop locations in Berlin
berlin <- metros[metro_names == "Berlin", ]

berlin_raster <- terra::crop(result, berlin) 

berlin_raster <- berlin_raster > 9

berlin_raster[berlin_raster == 0] = NA

mapviewOptions(fgb = FALSE)

mapview(raster::raster(berlin_raster), col.regions = c(NA, "darkgreen"),
            na.color = "transparent", legend = TRUE, map.type = "OpenStreetMap")
```

For instance, a score greater than 9 might be a suitable threshold indicating raster cells where a bike shop could be placed

## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/1PoOLfWxiAU")`

<details>
<summary> Meeting chat log </summary>

```
Cohort1 reviewed Transportation, Geomarketing, and Conclusion together 

00:47:35    Olivier Leroy:    And all the NA_ 😛
00:50:37    Olivier Leroy:    Dusseldorf is large!
00:51:59    Derek Sollberger (he/his):    The German regions are nuts!
https://en.wikipedia.org/wiki/NUTS_statistical_regions_of_Germany
00:52:05    Olivier Leroy:    OSM is so good in Germany, so many mappers
00:52:31    Derek Sollberger (he/his):    Reacted to "OSM is so good in Ge..." with 😻
00:55:18    Olivier Leroy:    Raster here save a lot of calculation
00:57:38    Olivier Leroy:    H3 indexes also
00:59:54    Tony Vota:    I have used the gistackexchange site with success
01:00:00    Olivier Leroy:    Reacted to "I have used the gist..." with 👍
01:08:59    Olivier Leroy:    end
```
</details>
